Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-09-27 Grassmannian_cells.tex

%\begin{lemma}
%  For any $m\ge1$, we have 
%  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0, \quad\dim\Ext(P_{m+1},P_m)=0.\]  
%  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
%\end{lemma}
%\begin{proof} 
%  Note that we have an Auslander-Reiten sequence
%  \begin{equation}
%    \label{eq:AR sequence}
%    0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0
%  \end{equation}
%  with irreducible homomorphisms $f_i$ which are also linearly independent. 
%  Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. 
%  If we pick $k<n$ linearly independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. 
%  Indeed otherwise, at least one of the homomorphisms would be forced to factor through a representation $Z\neq P_{m}$.
%\end{proof}
%\begin{remark}
%  Fix a basis $\{f_{m,1},\ldots,f_{m,n}\}$ for $\Hom(P_m,P_{m+1})$, $m\ge2$.
%  Define a basis $\{f_{\ell,1},\ldots,f_{\ell,n}\}$ for each $\Hom(P_\ell,P_{\ell+1})$, $1\le\ell\le m-1$, by downward induction as follows.  
%  Write $\iota_\ell:P_\ell\to P_{\ell+1}^n$ for the left-hand morphism from the sequence \eqref{eq:AR sequence} computed using $f_k=f_{\ell+1,k}$, $1\le k\le n$. 
%  Then the basis element $f_{\ell,k}$ is given by composing $\iota_\ell$ with the coordinate projection $\pi_k:P_{\ell+1}^n\to P_{\ell+1}$ onto the $k$-th factor.
%  Indeed, applying $\Hom(-,P_{\ell+1})$ to this sequence \eqref{eq:AR sequence} gives an isomorphism $\Hom(P_{\ell+1}^n,P_{\ell+1})\cong\Hom(P_\ell,P_{\ell+1})$ whereby the natural basis of coordinate projections on the left is taken to the basis $\{f_{\ell,k}\}$ on the right.
%\end{remark}
\begin{corollary}
  \label{cor:preprojective homomorphism duality}
  The sequence \eqref{eq:AR sequence} gives rise to an isomorphism $H_m^*\cong H_{m-1}$.
\end{corollary}
\begin{proof}
  Define a map $\Theta:H_m^*\to H_{m-1}$ by $\Theta(\varphi)=(id\overline{\otimes}\varphi)\circ\iota_{m-1}$, in words $\Theta(\varphi)$ acts on $x\in P_{m-1}$ by contracting with the second factor in $\iota_{m-1}(x)$ to give an element of $P_m$.
  Suppose $\varphi\in H_m^*$ is a nonzero functional on $H_m$ and let $V\subsetneq H_m$ denote the kernel of $\varphi$.
  Then $\Theta(\varphi)=0\in H_{m-1}$ if and only if the image of $\iota_{m-1}$ is contained in $P_m\otimes V\subsetneq P_m\otimes H_m$.
  But then Lemma~\ref{le:injective evaluation maps} implies $\iota_{m-1}$ is the zero map, a contradiction.
  Thus $\Theta:H_m^*\to H_{m-1}$ must be injective and hence an isomorphism.
\end{proof}
\begin{remark}
  The set $Gr(H_m)$ is naturally a poset and Corollary~\ref{cor:preprojective homomorphism duality} gives an identification of the opposite poset $Gr(H_m)^{op}\cong Gr(H_m^*)$ with $Gr(H_{m-1})$.
  We write $\bar{V}\subset H_{m-1}$ for the subspace corresponding to $V\subset H_m$ under this identification.
  Under the isomorphism of $H_{m-1}$ with $H_m^*$, we have $\bar{V}=(H_m/V)^*$.
\end{remark}

@@

\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).

>>>>

%\begin{lemma}
%  For any $m\ge1$, we have 
%  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0, \quad\dim\Ext(P_{m+1},P_m)=0.\]  
%  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
%\end{lemma}
%\begin{proof} 
%  Note that we have an Auslander-Reiten sequence
%  \begin{equation}
%    \label{eq:AR sequence}
%    0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0
%  \end{equation}
%  with irreducible homomorphisms $f_i$ which are also linearly independent. 
%  Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. 
%  If we pick $k<n$ linearly independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. 
%  Indeed otherwise, at least one of the homomorphisms would be forced to factor through a representation $Z\neq P_{m}$.
%\end{proof}
%\begin{remark}
%  Fix a basis $\{f_{m,1},\ldots,f_{m,n}\}$ for $\Hom(P_m,P_{m+1})$, $m\ge2$.
%  Define a basis $\{f_{\ell,1},\ldots,f_{\ell,n}\}$ for each $\Hom(P_\ell,P_{\ell+1})$, $1\le\ell\le m-1$, by downward induction as follows.  
%  Write $\iota_\ell:P_\ell\to P_{\ell+1}^n$ for the left-hand morphism from the sequence \eqref{eq:AR sequence} computed using $f_k=f_{\ell+1,k}$, $1\le k\le n$. 
%  Then the basis element $f_{\ell,k}$ is given by composing $\iota_\ell$ with the coordinate projection $\pi_k:P_{\ell+1}^n\to P_{\ell+1}$ onto the $k$-th factor.
%  Indeed, applying $\Hom(-,P_{\ell+1})$ to this sequence \eqref{eq:AR sequence} gives an isomorphism $\Hom(P_{\ell+1}^n,P_{\ell+1})\cong\Hom(P_\ell,P_{\ell+1})$ whereby the natural basis of coordinate projections on the left is taken to the basis $\{f_{\ell,k}\}$ on the right.
%\end{remark}
\begin{corollary}
  \label{cor:preprojective homomorphism duality}
  The sequence \eqref{eq:AR sequence} gives rise to an isomorphism $H_m^*\cong H_{m-1}$.
\end{corollary}
\begin{proof}
  Define a map $H_m^*\to H_{m-1}$ by $\varphi\mapsto\bar{\varphi}:=(id\overline{\otimes}\varphi)\circ\iota_{m-1}$, in words $\bar{\varphi}$ acts on $x\in P_{m-1}$ by contracting with the second factor in $\iota_{m-1}(x)$ to give an element of $P_m$.
  Suppose $\varphi\in H_m^*$ is a nonzero functional on $H_m$ and let $V\subsetneq H_m$ denote the kernel of $\varphi$.
  Then $\bar{\varphi}=0\in H_{m-1}$ if and only if the image of $\iota_{m-1}$ is contained in $P_m\otimes V\subsetneq P_m\otimes H_m$.
  But then Lemma~\ref{le:injective evaluation maps} implies $\iota_{m-1}$ is the zero map, a contradiction.
  Thus the map $H_m^*\to H_{m-1}$, $\varphi\mapsto\bar{\varphi}$ must be injective and hence an isomorphism.
\begin{remark}
  The set $Gr(H_m)$ is naturally a poset and Corollary~\ref{cor:preprojective homomorphism duality} gives an identification of the opposite poset $Gr(H_m)^{op}\cong Gr(H_m^*)$ with $Gr(H_{m-1})$.
  We write $\bar{V}\subset H_{m-1}$ for the subspace corresponding to $V\subset H_m$ under this identification.
  Under the isomorphism of $H_{m-1}$ with $H_m^*$, we have $\bar{V}=(H_m/V)^*$.
Following Remark~\ref{rem:reflection recursion}, the reflection functor $\Sigma_1$ provides a natural identification of $H_m$ with $H_{m-1}$ for $m\ge2$. 
Combining with Corollary~\ref{cor:preprojective homomorphism duality} we obtain a natural isomorphism $H_m^*\cong H_m$ for $m\ge2$.
The next observation is essential for the results to follow.
\begin{proposition}
  For $V\in Gr(H_m)$, $m\ge1$, we have $\bar{\bar{V}}=\tau V$.
\end{proposition}
\begin{proof}
  We show that $\bar{\bar{\varphi}}=\tau\varphi$ for $\varphi\in H_m$.
  Indeed, we have
  \[\bar{\bar{\varphi}}=(id\overline{\otimes}\bar{\varphi})\circ\iota_{m-2}=\Big(id\overline{\otimes}\big((id\overline{\otimes}\varphi)\circ\iota_{m-1}\big)\Big)\circ\iota_{m-2}.\]

@@

\begin{thebibliography}{10}
\bibitem{ass}
Assem, I., Simson, D., Skowronski, A.: Elements of the Representation Theory of Associative Algebras. Cambridge University Press, Cambridge 2007.
\bibitem{bb} Bialynicki-Birula, A.: Some theorems on actions of algebraic groups. Annals of Mathematics \textbf{98}, 480-497 (1973).
\bibitem{bgp}
Bernstein, I.~N., Gelfand, I.~M., Ponomarev, V.~A.: Coxeter functors, and Gabriel's theorem. Russian Mathematical Surveys \textbf{28}(2), 17-32 (1973).
\bibitem{brenner-butler} S. Benner, M.~C.~R. Butler, The equivalence of certain functors occuring in the representation theory of artin algebras and species, J. London Math. Soc., 14 (1976), 183-187.
\bibitem{cc}
Caldero, P., Chapoton, F.: Cluster algebras as {H}all algebras of quiver representations.
Commentarii Mathematici Helvetici \textbf{81}(3), 595-616 (2006).
\bibitem{ck}
  P. Caldero, B. Keller: From triangulated categories to cluster algebras II.  Ann. Sci. \'Ecole Norm. Sup. (4) \textbf{39} (2006), no. 6, pp.~983--1009.
\bibitem{cr}
Caldero, P., Reineke, M.: On the quiver Grassmannian in the acyclic case.
Journal of Pure and Applied Algebra \textbf{212}(11), 2369-2380 (2008).
\bibitem{cj} Crawley-Boevey, W., Jensen, Bernt Tore: A note on sub-bundles of vector bundles. Glasgow Mathematical Journal \textbf{48}, 459-462 (2006).
